A sheaf--function dictionary for algebraic tori over local fields

### 2013-10-01 shanghai.tex
Removed section 9, corrected and vastly simplified the proofs of Prop 9.1 (now Lem 8.3) and Cor 9.2 (now Prop 8.4)

While we have not yet fully explored the applications of such a geometrization, we want to highlight
some of the potential benefits of Theorem~\ref{thm:application}. 
%
Characters on $T$ have limited
functoriality: they can be pulled back along maps to $T$, but it is less clear what to do with them
in the presence of maps from $T$ to other groups.  Sheaves, on the other hand, come
equipped with a host of functors, including pushforward. 
We hope to use this to our advantage as we work toward an adaptation of Lusztig's theory of characters sheaves to the study of characters of admissible representations of reductive $p$-adic groups that extends beyond the case of $p$-adic tori.

DeBacker and Reeder have recently applied Deligne-Lusztig representations to part of the
local Langlands correspondence. % \cite{reeder-debacker:09a}.  
After developing Deligne-Lusztig representations, 
the next step Lusztig took in studying representations
of finite groups of Lie type was to introduce character sheaves for reductive groups over $\Fq$.
We hope that our quasicharacter sheaves 
can provide a stepping stone to a similar theory for
representations of $p$-adic groups.

The scope of the Langlands program has grown explosively in the past several decades, and
different parts of it apply to different kinds of representations.  In particular, the classic local Langlands
correspondence addresses complex (or $\ell$-adic) representations of real and $p$-adic groups, while
the geometric Langlands program aims at representations of groups in characteristic $p$.  The fact that
characteristic $p$ group schemes arise as the space underlying quasicharacter sheaves, regardless of
the characteristic of $K$, provides a potential bridge allowing the methods of the geometric Langlands
program to be applied to the classic context.

Finally, we are tantalized by the possibility of providing an alternate perspective on local class field theory
through quasicharacter sheaves.  The local Langlands correspondence for tori provides an isomorphism
\[
\Hh^1(\Weil{K}, \hat{T}_\ell) \cong \Hom(T(K), \EEx).
\]
A geometric description of the resulting isomorphism $\Hh^1(\Weil{K}, \hat{T}_\ell) \cong \QCiso{\GN{T}}$
would yield a different description of the Artin reciprocity map.

@@

We close the introduction with a more detailed description of the structure of this paper.
In Section~\ref{sec:category} we define quasicharacter sheaves and give an interpretation of
them as Weil sheaves with a Galois action.
Section~\ref{sec:Frob} introduces the trace of Frobenius homomorphism,
one direction of the sheaf--function dictionary.
We define pullbacks and products of quasicharacter sheaves in Section~\ref{sec:pullback}, and use
them to prove a functoriality result on the trace of Frobenius.
In Section~\ref{sec:basechange} we study how quasicharacter sheaves behave under base change
and Weil restriction.
Sections~\ref{sec:bounded} and \ref{sec:finite} describe bounded and finite quasicharacter sheaves,
two subcategories of $\QC(G)$.  These alternate descriptions of quasicharacter sheaves play a key
role in the proof of Proposition~\ref{prop:restriction}.
In section \ref{sec:connected} we show that bounded and finite quasicharacter sheaves are the same
as standard quasicharacter sheaves when $G$ is connected, and that trace of Frobenius is an isomorphism.
We shift to \'etale group schemes $G$ in Section~\ref{sec:etale} and give a more concrete interpretation
of quasicharacter sheaves in this case.  Using a notion of global sections for quasicharacter sheaves,
we give an isomorphism between $\QCiso{G}$ and $\Hh^1(\Weil{\Fq}, \Hom(G(\bFq), \EEx))$.
Section~\ref{sec:etalecohom} proves a result in Galois cohomology that shows that trace of Frobenius is
an isomorphism for \'etale group schemes.
In Section~\ref{sec:restriction} we show that for non-connected $G$, restriction to the identity component
induces an essentially surjective functor of quasicharacter sheaves.
Section~\ref{sec:snake} closes out Part \ref{part1} with the proof of the main theorem: trace of Frobenius
is an isomorphism for $G$ as long as the geometric component group is finitely generated.

@@

Now, let $\qcs{L} \ceq (\gqcs{L},\mu,\phi)$ be a quasicharacter sheaf on $G$. 
Set $\qcs{L}_1 \ceq q_1^* \qcs{L}$ and $\qcs{L}_2 \ceq q_2^* \qcs{L}$;
by Proposition~\ref{prop:pullback}, $\qcs{L}_1$ a quasicharacter sheaf on $G_1$ 
and $\qcs{L}_2$ is a quasicharacter sheaf on $G_2$.
We will obtain an isomorphism $\qcs{L} \iso  \qcs{L}_1\boxtimes \qcs{L}_2$.
%
Applying the functor $f^*$ to the isomorphism $\mu$ yields
\begin{equation}\label{eq:fm}
f^*\mu : f^* m^* \gqcs{L} \to f^*(\gqcs{L}\boxtimes \gqcs{L}) .
\end{equation}
We have already observed that $m\circ f = \id_G$, so $f^* m^* \gqcs{L} = \gqcs{L}$.
On the other hand, $f^*(\gqcs{L}\boxtimes \gqcs{L}) = f^*p_1^*\gqcs{L}\otimes f^* p_2^*\gqcs{L}$, 
by definition, and $f^*p_1^*\gqcs{L} = r_1^* q_1^* \gqcs{L}$ and $f^*p_2^*\gqcs{L} = r_2^* q_2^* \gqcs{L}$ follows from the properties of these maps, articulated above, 
so $f^*(\gqcs{L}\boxtimes \gqcs{L}) = \gqcs{L}_1\boxtimes \gqcs{L}_2$.
It follows that \eqref{eq:fm} gives an isomorphism $\gqcs{L} \to  \gqcs{L}_1\boxtimes \gqcs{L}_2$.
It is a matter of routine to show that this morphism satisfies Conditions~\ref{CS.2} as they apply here,
from which it follows that we have exhibited an isomorphism 
$\qcs{L} \to \qcs{L}_1\boxtimes \qcs{L}_2$ of quasicharacters sheaves on $G\times G$.
This completes the proof that $\boxtimes$ is essentially surjective.
\end{proof}

@@

\section{On the cohomology of the dual of an \'etale group scheme} \label{sec:etalecohom}

Although the next proposition makes no mention quasicharacter sheaves, 
it has immediate application to the problem of finding all quasicharacter sheaves
on finitely generated etale group schemes, as we will see at the end of the section.

\begin{proposition}\label{prop:X}
  Let $X$ be a finitely generated abelian group equipped with an
  action of $\Gal(\bFq/\Fq)$, continuous for the discrete topology on
  $X$.  Then
  \begin{align*}
    \Hh^1(\Fq, X^*) &\xrightarrow{\rho} \Hh^0(\Fq, X)^*, \\
    \rho([z])(x) &\ceq z(\Frob{\Fq})(x),
  \end{align*}
  is an isomorphism of groups.
\end{proposition}
\begin{proof}
  We first prove that $\rho$ is surjective.  Suppose that
  $f \in \Hom(X^{\Frob{\Fq}}, \EEx).$ Since $\EEx$ is divisible
  as an abelian group it is injective as a $\ZZ$-module and thus
  $\Ext^1_{\ZZ}(X/X^{\Frob{\Fq}}, \EEx) = 0$.  The long exact 
  sequence obtained by applying the functor
  $\Hom(\mbox{---}, \EEx)$ to
  \[
  \begin{tikzcd}
  0 \rar & X^{\Frob{\Fq}} \rar & X \rar & X/X^{\Frob{\Fq}} \rar & 0
  \end{tikzcd}
  \]
  gives the existence of an $\tilde{f} \in \Hom(X, \EEx)$ restricting to
  $f$.  Define a $1$-cocycle $z \in \Hh^1(\Fq, \Hom(X,\EEx))$ by
  mapping $\Frob{\Fq}$ to $\tilde{f}$.  Then $\rho([z]) = f$.

  Now suppose that $[z] \in \Hh^1(\Fq,\Hom(X,\EEx))$ with
  $\rho([z]) = 1$; write $\tilde{f}$ for $z(\Frob{\Fq})$.  The injectivity
  of $\rho$ is equivalent to the existence of a homomorphism $\psi
  \colon X \rightarrow \EEx$ with
  \[
  \tilde{f}(x) = \frac{\psi(\Frob{\Fq}(x))}{\psi(x)}
  \]
  for $x \in X$.  In fact, we may replace the construction of $\psi$
    \Clifton{I'm not sure I understand what you mean here.}
  for all $\ZZ[\Gal(\bFq/\Fq)]$-modules $X$ with the construction of
  analogous $\psi$s for a different class of modules.

  The $\ZZ[\Gal(\bFq/\Fq)]$-module $X$ decomposes into 
  a finite direct sum of cyclic $\ZZ[\Gal(\bFq/\Fq)]$-modules. 
  To see this, observe first that since the action of $\Gal(\bFq/\Fq)$ is continuous, 
  there is a positive integer $m$ so that $\Frob{\Fq}^m$ acts trivially on $X$; 
  write $\Fqm$ for the degree $m$ extension of $\Fq$ in $\bFq$ 
  and $\Gamma_m$ for $\Gal(\Fqm/\Fq)$.  Since
  $T^m-1$ factors as a squarefree product of cyclotomic polynomials,
  the Chinese remainder theorem expresses $\ZZ[\Gamma_m]$ as a product
  of $\ZZ[\zeta_d]$ for $d$ dividing $m$.  Each $\ZZ[\zeta_d]$ is the
  maximal order in a number field and thus $\ZZ[\Gamma_m]$ is a
  product of Dedekind domains. Combining \cite{Warfield:Decomposability}*{Theorem 3}
  and \cite{brandal:CommRingsFGModDecompose}*{Lemma 4.1}, it now follows that 
  $X$ decomposes into a direct sum of cyclic $\ZZ[\Gamma_m]$-modules, 
  each of the form $\ZZ[\Gamma_m] / I$ for some ideal $I \subset \ZZ[\Gamma_m]$.
  The ideal $I$ corresponds to a direct product of ideals $I_d$ in the
  $\ZZ[\zeta_d]$, and within each $\ZZ[\zeta_d]$ the Chinese remainder
  theorem allows a further decomposition of $\ZZ[\zeta_d] / I_d$ as a
  direct sum of $\ZZ[\zeta_d]$-modules of the form $\ZZ[\zeta_d]/P^s$.
  So if we can construct, for every cyclotomic ring $\ZZ[\zeta_d]$,
  prime ideal $P \subset \ZZ[\zeta_d]$ and positive integer $s$, a
  $\psi \colon \ZZ[\zeta_d]/P^s \rightarrow \EEx$ with
  \begin{equation} \label{eq:psi-condition}
    \tilde{f}(x) = \frac{\psi(\zeta_d \cdot x)}{\psi(x)},
  \end{equation}
  then the existence of the original $\psi$ follows because $X$ is a
  direct sum of such $\ZZ[\zeta_d]/P^s$.

  We now fix a $d$, $P$ and $s$, write $\zeta$ for $\zeta_d$ and
  choose a polynomial
  \[
  h(T) \ceq a_rT^r + a_{r-1}T^{r-1} + \cdots + a_0
  \]
  so that $P^s = (p^s, h(\zeta_d))$.
  % Setting $\psi = 1$ will satisfy Equation \ref{eq:psi-condition}
  % when $d = 1$.
  An an abelian group, the quotient $\ZZ[\zeta] / P^s$ is generated by
  $\zeta^i$ for $0 \le i < \phi(d)$; we define
  \begin{equation} \label{eq:psi-def}
    \psi(\zeta^i) \ceq \alpha \prod_{j=0}^{i-1} \tilde{f}(\zeta^j)
  \end{equation}
  for an $\alpha \in \EEx$ to be chosen later.  Since there are
  relations among the $\zeta^i$ we first need to check that $\psi$ is
  well-defined.

  The relations among the generators $\zeta^i$ are additively
  generated by $\zeta^ch(\zeta)$ as $c$ ranges from $0$ to
  $\phi(d)-r-1$.  If we set
  \[
  A_c \ceq \sum_{i=0}^r a_i \sum_{j=0}^{i+c-1} \zeta^j
  \]
  then
  \[
  \psi(\zeta^ch(\zeta)) = \alpha^{h(1)} \tilde{f}(A_c).
  \]
  Now consider $(\zeta - 1)A_c$: the sums telescope and we get that
  $(\zeta-1)A_c = \zeta^ch(\zeta) - h(1).$ There are now two
  possibilities:
  \begin{enumerate}
  \item $d$ is a power of $p$; write $s = \phi(d)m + n$.  We may take
    $h(T)$ to be $p^m(T-1)^n$, in which case $\zeta^ch(\zeta) - h(1) \in P^s$,
    so $A_c$ is fixed under multiplication by $\zeta$ in
    $\ZZ[\zeta]/P^s$ and thus $\tilde{f}(A_c) = 1$.  In this case
    $\alpha^{h(1)} = 1$ regardless of $\alpha$.
  \item $d$ is not a power of $p$.  If two distinct primes divide $d$
    then $\zeta - 1$ is a unit; otherwise $\zeta - 1$ is a unit modulo
    $P^s$.  Let $\omega \in \ZZ[\zeta]$ satisfy
    $\omega(\zeta - 1) \equiv 1 \pmod{P^s}$ and set
    %\begin{equation} \label{alpha-def}
     $ \alpha \ceq \tilde{f}(\omega)$.
    %\end{equation}
    Then $A_c + h(1)\omega$ is fixed under multiplication by $\zeta$
    and thus $\tilde{f}(A_c) = \tilde{f}(-h(1)\omega)$.
  \end{enumerate}
  In either case, $\psi(\zeta^ch(\zeta)) = 1$ for all $c$.  The only
  other relations imposed on the $\zeta^i$ are that they are all
  killed under multiplication by $p^s$.  Since $\tilde{f}$ is also a
  homomorphism $\ZZ[\zeta]/P^s \rightarrow \EEx$ the values of
  $\tilde{f}$ and thus of $\psi$ will be $p^s$-roots of unity.
  Therefore $\psi$ is well defined.

  We now check Equation~\eqref{eq:psi-condition} for
  $x \ceq \zeta^{\phi(d)-1}$.  Write
  \[\Phi_d(T) \ceq T^{\phi(d)} - b_{\phi(d)-1}T^{\phi(d)-1} - \cdots - b_0\]
  and set
  \[
  B \ceq b_i \sum_{j=0}^{i-1} \zeta^j -  \sum_{i=0}^{\phi(d)-1} \zeta^i.
  \]
  We have

  \begin{align*}
    \frac{\psi(\zeta^{\phi(d)})}{\psi(\zeta^{\phi(d)-1})\tilde{f}(\zeta^{\phi(d)-1})}
    &= \frac{\psi(b_{\phi(d)-1}\zeta^{\phi(d)-1} + \cdots + b_0)}{\psi(\zeta^{\phi(d)-1})\tilde{f}(\zeta^{\phi(d)-1})} 
    = \alpha^{-\Phi_d(1)} \tilde{f}(B).
  \end{align*}
  As before we see that $(\zeta - 1)B = \Phi_d(1) - \Phi_d(\zeta) = \Phi_d(1)$.
  We consider the same cases as before:

  \begin{enumerate}
  \item $d$ is a power of $p$, in which case $\Phi_d(1) = p$.  Thus
    $(\zeta-1)B = p = u(\zeta-1)^{\phi(d)}$ for a unit $u \in \ZZ[\zeta]$
    and $B - u(\zeta-1)^{\phi(d)-1}$ is fixed by $\zeta$,
    so that $\tilde{f}(B) = \tilde{f}(u(\zeta - 1)^{\phi(d)-1})$.  The
    multiplicative order of $\tilde{f}(u(\zeta-1)^{\phi(d)-1})$ is at most
    $\lceil \frac{s - 1 - (\phi(d)-1)}{\phi(d)}\rceil = \lceil \frac{s}{\phi(d)} \rceil - 1$
    while the additive order of $1 \in \ZZ[\zeta]/P^s$ is $\lceil \frac{s}{\phi(d)} \rceil$.
    Thus we may choose $\alpha = \psi(1)$ with $\alpha^p = \tilde{f}(B)$.
  \item $d$ is not a power of $p$ and $\omega$ is the inverse of
    $\zeta-1$ modulo $P^s$.  Then $\tilde{f}(B) = \tilde{f}(\omega\Phi_d(1))$.
  \end{enumerate}
  In either case, $\alpha^{-\Phi_d(1)}\tilde{f}(B) = 1$ and thus
  $\psi$ satisfies Equation~\eqref{eq:psi-condition} for
  $x \ceq \zeta^{\phi(d)-1}$.  The definition of $\psi$ in \ref{eq:psi-def}
  immediately implies that $\psi$ satisfies equation
  \ref{eq:psi-condition} for lesser powers of $\zeta$, and thus for
  all $x$ by linearity.  The existence of $\psi$ shows that $\rho$ is injective.
\end{proof}

\begin{corollary}
 If $G$ is a commutative, \'etale group scheme over $\Fq$ and $G(\bFq)$ is finitely
 generated then
  \[
  \trFrob{G} : \QCiso{G} \to G(\Fq)^*
  \]
  is an isomorphism of groups.
\end{corollary}
\begin{proof}
This follows from Proposition~\ref{prop:etale} by setting $X=G(\bFq)$ in Proposition~\ref{prop:X}.
Recall that $G \mapsto G(\bFq)$ defines an equivalence of categories from commutative \'etale group schemes over $\Fq$ to commutative groups equipped with \emph{continuous} actions by $\Gal(\bFq/\Fq)$; see, for example, \cite{vdGeer-Moonen:AbelianVarieties}*{Prop 3.26}.
\end{proof}

@@

  In the proof of Proposition~\ref{prop:restriction} we saw that $G^0$ is an algebraic group over $\Fq$. % clearly connected. 
  It now follows from Proposition~\ref{prop:connected} 
  that $\ker \trFrob{G^0} =0$ and $\coker \trFrob{G^0}=0$. 
  On the other hand, it follows from Propositions~\ref{prop:etale} and \ref{prop:X}, $\ker \trFrob{\pi_0(G)}=0$
  and $\coker \trFrob{\pi_0(G)}=0$.

@@

  The sequence of groups of isomorphism classes of quasicharacter sheaves is right-exact: it is exact at
  $\QCiso{G^0}$ by Proposition~\ref{prop:restriction}; and
  it is exact at $\QCiso{G}$ because the left-square
  commutes (by Proposition~\ref{prop:pullback}), the bottom sequence
  is exact at $G(\Fq)^*$ and $\coker \trFrob{\pi_0(G)} =0$
  (by Propositions~\ref{prop:etale} and \ref{prop:X}) and
  \[
  \image(\QCiso{\pi_0(G)} \to \QCiso{G}) \subseteq \ker(\QCiso{G} \to \QCiso{G^0}),
  \]
  by the last part of Proposition~\ref{prop:pullback}. Specifically,
  suppose $\qcs{L}$ is a quasicharacter sheaf on $G$ that is trivial on
  $G^0$. Then $\trFrob{\qcs{L}}\vert_{G^0(\Fq)} = \trFrob{\qcs{L}^0} =0$,
  by Proposition~\ref{prop:pullback} again. Since the sequence of
  character groups is exact at $G(\Fq)^*$, there is
  some $\chi \in \pi_0(G)(\Fq)^*$ such that the
  pull-back of $\chi$ along $G(\Fq)\to \pi_0(G)(\Fq)$ is
  $\trFrob{\qcs{L}}$. By Propositions~\ref{prop:etale} and \ref{prop:X},
  there is a a quasicharacter sheaf $\qcs{L}_\chi$ on $\pi_0(G)$, unique up
  to isomorphism, such that $\trFrob{\qcs{L}_\chi} = \chi$. By
  Proposition~\ref{prop:pullback} again, $\qcs{L}$ is isomorphic to the
  image of $\qcs{L}_\chi$ under the monoidal functor $\QC(\pi_0(G)) \to \QC(G)$.

>>>>

While we have not yet fully explored the applications of such a geometrization, we want to highlight
some of the potential benefits of Theorem~\ref{thm:application}. 
%\begin{enumerate}
%\item
Characters on $T$ have limited
functoriality: they can be pulled back along maps to $T$, but it is less clear what to do with them
in the presence of maps from $T$ to other groups.  Sheaves, on the other hand, come
equipped with a host of functors, including pushforward. 
We hope to use this to our advantage as we work toward an adaptation of Lusztig's theory of characters sheaves to the study of characters of admissible representations of reductive $p$-adic groups that extends beyond the case of $p$-adic tori. 
%\item
DeBacker and Reeder have recently applied Deligne-Lusztig representations to part of the
local Langlands correspondence. % \cite{reeder-debacker:09a}.  
After developing Deligne-Lusztig representations, 
the next step Lusztig took in studying representations
of finite groups of Lie type was to introduce character sheaves for reductive groups over $\Fq$.
We hope that our quasicharacter sheaves 
can provide a stepping stone to a similar theory for
representations of $p$-adic groups. 
%\item
The scope of the Langlands program has grown explosively in the past several decades, and
different parts of it apply to different kinds of representations.  In particular, the classic local Langlands
correspondence addresses complex (or $\ell$-adic) representations of real and $p$-adic groups, while
the geometric Langlands program aims at representations of groups in characteristic $p$.  The fact that
characteristic $p$ group schemes arise as the space underlying quasicharacter sheaves, regardless of
the characteristic of $K$, provides a potential bridge allowing the methods of the geometric Langlands
program to be applied to the classic context. 
%\item
Finally, we are tantalized by the possibility of providing an alternate perspective on local class field theory
through quasicharacter sheaves.  The local Langlands correspondence for tori provides an isomorphism
\[
\Hh^1(\Weil{K}, \hat{T}_\ell) \cong \Hom(T(K), \EEx).
\]
A geometric description of the resulting isomorphism $\Hh^1(\Weil{K}, \hat{T}_\ell) \cong \QCiso{\GN{T}}$
would yield a different description of the Artin reciprocity map.
%\end{enumerate}

@@

We close the introduction with a more detailed description of the structure of this paper.
In Section~\ref{sec:category} we define quasicharacter sheaves and give an interpretation of
them as Weil sheaves with a Galois action.
Section~\ref{sec:Frob} introduces the trace of Frobenius homomorphism,
one direction of the sheaf--function dictionary.
We define pullbacks and products of quasicharacter sheaves in Section~\ref{sec:pullback}, and use
them to prove a functoriality result on the trace of Frobenius.
In Section~\ref{sec:basechange} we study how quasicharacter sheaves behave under base change
and Weil restriction.
Sections~\ref{sec:bounded} and \ref{sec:finite} describe bounded and finite quasicharacter sheaves,
two subcategories of $\QC(G)$.  These alternate descriptions of quasicharacter sheaves play a key
role in the proof of Proposition~\ref{prop:restriction}.
In section \ref{sec:connected} we show that bounded and finite quasicharacter sheaves are the same
as standard quasicharacter sheaves when $G$ is connected, and that trace of Frobenius is an isomorphism.
We shift to \'etale group schemes $G$ in Section~\ref{sec:etale} and give a more concrete interpretation
of quasicharacter sheaves in this case.  Using a notion of global sections for quasicharacter sheaves,
we give isomorphisms between $\QCiso{G}$, $\Hh^1(\Weil{\Fq}, \Hom(G(\bFq), \EEx))$ and $\Hom(G(\Fq), \EEx)$.
In Section~\ref{sec:restriction} we show that for non-connected $G$, restriction to the identity component
induces an essentially surjective functor of quasicharacter sheaves.
Section~\ref{sec:snake} closes out Part \ref{part1} with the proof of the main theorem: trace of Frobenius
is an isomorphism for $G$ as long as the geometric component group is finitely generated.

@@

Now, let $\qcs{L} = (\gqcs{L},\mu,\phi)$ be a quasicharacter sheaf on $G$. 
Set $\qcs{L}_1 \ceq q_1^* \qcs{L}$ and $\qcs{L}_2 \ceq q_2^* \qcs{L}$;
by Proposition~\ref{prop:pullback}, $\qcs{L}_1$ a quasicharacter sheaf on $G_1$ 
and $\qcs{L}_2$ is a quasicharacter sheaf on $G_2$.
We will obtain an isomorphism $\qcs{L} \iso  \qcs{L}_1\boxtimes \qcs{L}_2$.
%
Applying the functor $f^*$ to the isomorphism $\mu$ yields the isomorphism
\begin{equation}\label{eq:fm}
f^*\mu : f^* m^* \gqcs{L} \to f^*(\gqcs{L}\boxtimes \gqcs{L}) .
\end{equation}
We have already observed that $m\circ f = \id_G$, so $f^* m^* \gqcs{L} = \gqcs{L}$ (canonical isomorphism).
On the other hand, $f^*(\gqcs{L}\boxtimes \gqcs{L}) = f^*p_1^*\gqcs{L}\otimes f^* p_2^*\gqcs{L}$, 
by definition, and $f^*p_1^*\gqcs{L} = r_1^* q_1^* \gqcs{L}$ and $f^*p_2^*\gqcs{L} = r_2^* q_2^* \gqcs{L}$ follows from the properties of these maps, articulated above, 
so $f^*(\gqcs{L}\boxtimes \gqcs{L}) = \gqcs{L}_1\boxtimes \gqcs{L}_2$.
It follows that \eqref{eq:fm} gives an isomorphism $\gqcs{L} \to  \gqcs{L}_1\boxtimes \gqcs{L}_2$.
It is a matter of routine to show that this morphism satisfies Conditions~\ref{CS.2} as they apply here,
from which it follows that we have exhibited an isomorphism 
$\qcs{L} \to \qcs{L}_1\boxtimes \qcs{L}_2$ of quasicharacters sheaves on $G\times G$.
This completes the proof that $\boxtimes$ is essentially surjective.
\end{proof}

@@

In order to relate $\QCiso{G}$ to $G(k)^*$, we need a simple result relating duals, invariants and coinvariants.

\begin{lemma} \label{lem:dual-inv}
Let $X$ be a finitely generated abelian group equipped with an action of $\Weil{\Fq}$, and write $\Frob{} = \Frob{\Fq}$.  Then
\begin{align*}
 (X^*)_{\Frob{}} &\to (X^{\Frob{}})^* \\
 [f] &\mapsto f|_{X^{\Frob{}}}
\end{align*}
is an isomorphism.
\end{lemma}
\begin{proof}
We can describe $X^{\Frob{}}$ as the kernel of the map $X \xrightarrow{\Frob{}-1} X$;
let $Y = (\Frob{}-1)X$ be the augmentation ideal.  Dualizing the sequence
\[
 0 \to X^{\Frob{}} \to X \to Y \to 0
\]
yields
\[
 0 \to Y^* \to X^* \to (X^{\Frob{}})^* \to \Ext^1_\ZZ(Y, \EEx).
\]
Since $\Ext^1_\ZZ(-,\EEx)$ vanishes, we get a natural isomorphism from the cokernel of $Y^* \xrightarrow{\Frob{}-1} X^*$ to $(X^{\Frob{}})^*$.
\end{proof}

\begin{proposition} \label{prop:etale-iso}
 If $G$ is a commutative, \'etale group scheme over $\Fq$ and $G(\bFq)$ is finitely
 generated then
 \[
  \trFrob{G} : \QCiso{G} \to G(\Fq)^*
 \]
 is an isomorphism of groups.
\end{proposition}
\begin{proof}
There is a natural isomorphism $\Hh^1(\Weil{\Fq}, G(\bFq)^*) \to (G(\bFq)^*)_{\Frob{}}$ defined by $[z] \mapsto z(\Frob{})$.  The proposition now follows directly from Proposition~\ref{prop:etale} and Lemma~\ref{lem:dual-inv}.
\end{proof}

@@

  In the proof of Proposition~\ref{prop:restriction} we saw that $G^0$ is an algebraic group over $\Fq$. % clearly connected. 
  It now follows from Proposition~\ref{prop:connected} 
  that $\ker \trFrob{G^0} =0$ and $\coker \trFrob{G^0}=0$. 
  On the other hand, it follows from Propositions~\ref{prop:etale-iso} that $\ker \trFrob{\pi_0(G)}=0$
  and $\coker \trFrob{\pi_0(G)}=0$.

@@

  The sequence of groups of isomorphism classes of quasicharacter sheaves is right-exact: it is exact at
  $\QCiso{G^0}$ by Proposition~\ref{prop:restriction}; and
  it is exact at $\QCiso{G}$ because the left-square
  commutes (by Proposition~\ref{prop:pullback}), the bottom sequence
  is exact at $G(\Fq)^*$ and $\coker \trFrob{\pi_0(G)} =0$
  (by Proposition~\ref{prop:etale-iso}) and
  \[
  \image(\QCiso{\pi_0(G)} \to \QCiso{G}) \subseteq \ker(\QCiso{G} \to \QCiso{G^0}),
  \]
  by the last part of Proposition~\ref{prop:pullback}. Specifically,
  suppose $\qcs{L}$ is a quasicharacter sheaf on $G$ that is trivial on
  $G^0$. Then $\trFrob{\qcs{L}}\vert_{G^0(\Fq)} = \trFrob{\qcs{L}^0} =0$,
  by Proposition~\ref{prop:pullback} again. Since the sequence of
  character groups is exact at $G(\Fq)^*$, there is
  some $\chi \in \pi_0(G)(\Fq)^*$ such that the
  pull-back of $\chi$ along $G(\Fq)\to \pi_0(G)(\Fq)$ is
  $\trFrob{\qcs{L}}$. By Proposition~\ref{prop:etale-iso},
  there is a a quasicharacter sheaf $\qcs{L}_\chi$ on $\pi_0(G)$, unique up
  to isomorphism, such that $\trFrob{\qcs{L}_\chi} = \chi$. By
  Proposition~\ref{prop:pullback} again, $\qcs{L}$ is isomorphic to the
  image of $\qcs{L}_\chi$ under the monoidal functor $\QC(\pi_0(G)) \to \QC(G)$.
